Data cleaning and processing script

### 2023-02-24 code/analysis_code/exploratory_analysis_LML.qmd
exploratory analysis

```{r}
here()
n1_n2_clean <- readRDS(here("data/processed_data/year2_cfx_clean.rds"))
plant_clean <- readRDS(here("data/processed_data/year2_plant_clean.rds"))
n1_n2_plant_combine <- readRDS(here("data/processed_data/year2_cfx_plant_combined.rds"))
covid_combine <- readRDS(here("data/processed_data/all_covid_combined.rds"))
all_data_combine <- readRDS(here("data/processed_data/year2_all_data_combined.rds"))
recovery_percent <- readRDS(here("data/processed_data/recovery_percentages.rds"))

@@

### Look at `all_data_combine` and determine visualization to make

@@

Some plots I want to generate initially are

1.  Case averages over time
2.  Total copies over time (N1 and N2)

### Look at case average over time

@@

### Look at log10 total copies in wastewater over time for N1

First, average total copies of biological replicates. Then, add total copies from each facility.

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(x = date,y = log10(sum_total_copies))) +
  geom_point(color = "red") +
  stat_smooth(method = "gam")

@@

### Look at log10 total copies in wastewater over time for N2

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>% 
  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Look at log10 copies per liter in N1

>>>>

```{r}
here()
n1_n2_clean <- readRDS(here("data/processed_data/year2_cfx_clean.rds"))
plant_clean <- readRDS(here("data/processed_data/year2_plant_clean.rds"))
n1_n2_plant_combine <- readRDS(here("data/processed_data/year2_cfx_plant_combined.rds"))
covid_combine <- readRDS(here("data/processed_data/all_covid_combined.rds"))
all_data_combine <- readRDS(here("data/processed_data/year2_all_data_combined.rds"))
recovery_percent <- readRDS(here("data/processed_data/recovery_percentages.rds"))
n1_n2n2_avg <- readRDS(here("data/processed_data/year2_cfx_avg.rds"))

@@

### Look at `all_data_combine`

@@

### Variation analysis

Graphs to make:

-   histogram of case count observations

-   histogram of 7 day moving average case observations

-   histogram of total daily copies

-   histogram of copies per L

-   histogram of percent positive qPCR assays

### Co-variation analysis

Graphs to make:

#### Histogram of clinical case 7 day moving average

```{r}
covid_combine %>% ggplot(aes(pcr_pos_7dma)) +
  geom_histogram()
summary(covid_combine$pcr_pos_7dma)
```

#### Histogram of PCR positive clinical cases

```{r}
covid_combine %>% ggplot(aes(pcr_pos)) +
  geom_histogram()
summary(covid_combine$pcr_pos)
```

#### Histogram of clinical case 14 day moving average

```{r}
covid_combine %>% ggplot(aes(pcr_pos_14dma)) +
  geom_histogram()
summary(covid_combine$pcr_pos_14dma)
```

#### Histogram of total daily viral copies at county level

```{r}
sum_daily_copies_county <- all_data_combine %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies))

  sum_daily_copies_county %>% ggplot(aes(log10(sum_total_copies))) +
  geom_histogram()

  log10_sum_daily_copies <- log10(sum_daily_copies_county$sum_total_copies)
  summary(log10_sum_daily_copies)
```

### Look at case 7 day moving average over time

@@

Case 7 day moving average overtime, bar plot and GAM model

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma),
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_7dma)) +
  geom_bar(stat = "identity", color = '#634490') +
  stat_smooth(method = "gam", color = 'black')

@@

PCR positives per day over time, bar plot and GAM model

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_pcr_pos)) +
  geom_bar(stat = "identity", color = '#634490') +
  stat_smooth(method = "gam", color = 'black') 
```

### Look at log10 total copies in wastewater over time (avg N1/N2)

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  #drop_na() %>% #remove NAs from analysis
  group_by(date, facility) %>% #this will allow me to average the two targets
  summarize(mean_total_copies = mean(total_copies)) %>% #average across N1 and N2
  ungroup(facility) %>% #need to do this so I can add the avg total copies of each WRF
  summarize(sum_total_copies = sum(mean_total_copies)) %>% #total daily county copies
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% 
  group_by(date, target, facility) %>% 
  drop_na(target) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies)) %>%

  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% filter(target == "N1") %>%

  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>%

  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

### Look at log10 copies per liter in N2

### 2023-02-24 code/processing_code/data_processing_LML.qmd
exploratory analysis

### Hypotheses so far

Waste water surveillance data will begin to correlate less with the clinical data as we progress further into the pandemic. (as rapid, at-home testing becomes more prevalent)

Catchment-level surveillance data will correlate better to waste water catchment-level case data than it will with county-level data because people are more likely to poop at their own house and SARS-CoV-2 particles are shed more often in feces than they are in urine (Kim et al. 2020).

@@

I will start by combining the CFX output data with the plant flow data and calling the dataframe `wbe`

@@

@@

```{r}
n1_n2_plant <- left_join(wbe_year2, plant_clean, by = c("date","facility")) %>% 
  mutate(total_copies = copy_num_L * influent_flow_L)

>>>>

## Questions I have so far

### Variation

1.  What is the distribution like of SARS-CoV-2 viral load observations? What does this distribution look like at catchment levels vs at the county level?
2.  What is the distribution like of N1 and N2 qPCR percent positive observations? What does this distribution look like at catchment levels vs at the county level?
3.  What is the distribution like of clinical case observations? What does this distribution look like at catchment levels vs at the county level?

### Covariation

1.  Is there a relationship between viral load and clinical case observations?

2.  Is there a relationship between qPCR percent positive and clinical case observations?

3.  What do viral load trends look like over time?

4.  What do clinical case trends look like over time?

5.  How do each of these temporal trends compare to one another over time?

@@

I will start by combining the CFX output data with the plant flow data and calling the data frame `wbe`

@@

### Summarizing CFX data for downstream processing

```{r}
wbe_year2_avg <- wbe_year2 %>% group_by(date,facility,target,collection_num) %>% 
  summarize(avg_copy_L = mean(copy_num_L), 
            avg_copy_uL_rxn = mean(copy_num_uL_rxn)) %>% 
  ungroup()
saveRDS(wbe_year2_avg, here("data/processed_data/year2_cfx_avg.rds"))
```

@@

```{r}
n1_n2_plant <- left_join(wbe_year2_avg, plant_clean, by = c("date","facility")) %>% 
  mutate(total_copies = avg_copy_L * influent_flow_L)
